```{r}
#test
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(MASS) # for hills data and rlm function
library(dplyr)
library(ggformula)

```

## Stores example
```{r}
set.seed(30)
n_cust = c(sample(10:20, 15, replace=T), sample(100:200, 15, replace=T)) # n_cust = number of customers per week at diff stores

avg_time = numeric(length = 30)
avg_spent = numeric(length = 30)
for(store in 1:30){
  time = runif(n_cust[store], 5, 40)
  spent = .5*time + 5 + rnorm(n_cust[store], 0, 5)
  avg_time[store] = mean(time)
  avg_spent[store] = mean(spent)
}

gf_point(avg_spent ~ avg_time)

```

```{r}
customers = ifelse(n_cust < 21, "10-20", "100-200")
gf_point(avg_spent ~ avg_time, color =~ customers, shape =~ customers)
```
True model (used to create the simulation) is $$\hat{y} = 5 + 0.5*time$$
```{r}
# Unweighted
fit1 = lm(avg_spent ~ avg_time)
summary(fit1)
```
```{r}
# Weighted
fit2 = lm(avg_spent ~ avg_time, weights = n_cust)
summary(fit2)
```

```{r}
lin_mod = data.frame(model = c("Unweighted", "Weighted", "True"), intercepts = c(fit1$coef[1], fit2$coef[1], 5), slopes = c(fit1$coef[2], fit2$coef[2], 0.5))

gf_point(avg_spent ~ avg_time) %>%
  gf_abline(slope = ~slopes, intercept = ~intercepts, 
            color =~ model, lty =~ model, lwd = 1.1, 
            data = lin_mod)
```
```{r}
set.seed(12)
x = runif(100)
y = 1 + 3*x + rnorm(100, 0, 1 + 3*x)
gf_point(y ~ x)
```
## IWLS
```{r}
fit.w = lm(y~x)

oldcoef = rep(0,length(fit.w$coef))
newcoef = fit.w$coef
iter = 0

while(sum(abs(oldcoef-newcoef)) > .0001 & iter < 100){
  w = 1/(fit.w$fitted.values^2)
  fit.w = lm(y~x, weights=w)
  
  iter = iter + 1
  oldcoef = newcoef
  newcoef = fit.w$coef
}

```

### A more complicated model for noise terms
```{r}
set.seed(12)
x = runif(100)
y = 3 + 2*x + rnorm(100, 0, sqrt(1-x)) # rnorm uses standard deviation instead of variance
gf_point(y ~ x)
```

```{r}
while(sum(abs(oldcoef-newcoef)) > .0001 & iter < 100){
  fit.epsilon = lm(fit.w$residuals^2 ~ x)
  w = 1/fit.epsilon$fitted.values
  fit.w = lm(y~x, weights=w)
  
  iter = iter + 1
  oldcoef = newcoef
  newcoef = fit.w$coef
}

```

### Hill racing
```{r}
fit_hills = lm(time ~ dist, data = hills)

hills %>%
  gf_point(time ~ dist) %>%
  gf_abline(intercept = fit_hills$coef[1], slope = fit_hills$coef[2]) %>%
  gf_label(195 ~ 22, label = "OLS")
```

```{r}
plot(fit_hills)
```

### While loop for Tukey's Bisquare method
```{r}
fit.w = lm(time ~ dist, data = hills)

oldcoef = rep(0,length(fit.w$coef))
newcoef = fit.w$coef
iter = 0 

while(sum(abs(oldcoef-newcoef)) > .0001 & iter < 100){
    MAR = median(abs(fit.w$residuals)) # Median absolute value of residuals
    sigma = MAR/0.6745
    k = 4.685*sigma
    sqrt_w = pmax(1-(fit.w$residuals/k)^2, 0)
    w = sqrt_w^2
    fit.w = lm(time ~ dist, data = hills, weights=w)

    iter = iter + 1
    oldcoef = newcoef
    newcoef = fit.w$coef
}

```

Why do we use pmax?
```{r}
tukey = data.frame(r = seq(-6, 6, length = 1000))
k = 4.685

tukey <- tukey %>%
  mutate(sqrt_wt = 1-(r/k)^2)  
```

```{r}
tukey %>%
  gf_line(sqrt_wt ~ r) %>%
  gf_hline(yintercept = 0, lty = 2)
```
Highlighting the higher of the two lines
```{r}
tukey <- tukey %>%
  mutate(selected = ifelse(abs(r)< k, sqrt_wt, 0))

tukey %>%
  gf_line(selected ~ r, lwd = 3, color = "yellow") %>%
  gf_line(sqrt_wt ~ r) %>%
  gf_hline(yintercept = 0, lty = 2)
```

### rlm
```{r}
fit_bisquare = rlm(time ~ dist, data = hills, psi = psi.bisquare)
fit_huber = rlm(time ~ dist, data = hills, psi = psi.huber)

rlm_mod = data.frame(model = c("OLS", "Tukey", "Huber"), 
                     intercepts = c(fit_hills$coef[1], fit_bisquare$coef[1], fit_huber$coef[1]), 
                     slopes = c(fit_hills$coef[2], fit_bisquare$coef[2], fit_huber$coef[2]))

hills %>%
  gf_point(time ~ dist) %>%
  gf_abline(slope = ~slopes, intercept = ~intercepts, 
            color =~ model, lty =~ model, lwd = 1.1, 
            data = rlm_mod)
```

### Plotting the weights

```{r}
hills2 <- hills %>%
  tibble::rownames_to_column("Race")

hills2 <- hills2 %>%
  mutate(my_label = case_when(fit_bisquare$w < .8 ~ Race,
                           TRUE ~ ""))

gf_point(fit_bisquare$w ~ 1:dim(hills)[1]) %>%
  gf_text(fit_bisquare$w - .02 ~ 1:dim(hills)[1], 
           label = hills2$my_label)
```

An alternative approach using base R graphics.  Run this code in the console or by knitting the R Markdown document (not by clicking the green arrow in the code cell).
```{r}
plot(fit_bisquare$w)
small_weights = which(fit_bisquare$w < .8)
car::showLabels(1:dim(hills)[1], fit_bisquare$w, 
           labels = rownames(hills),
           method = small_weights)
```

### Additional notes
```{r}
huber = rlm(calls~year, data=phones)
huber = rlm(calls~year, data = phones, maxit = 50)  
```

```{r}
set.seed(42)
x=c(runif(9),2)
y=2*x+rnorm(10,0,.25)
y[10]=1

bisquare = rlm(y ~ x, psi = psi.bisquare)
gf_point(y ~ x) %>%
  gf_abline(intercept = bisquare$coef[1], slope = bisquare$coef[2])
```
```{r}
bisquare2 = rlm(y~x, psi=psi.bisquare,
  weights=c(1,1,1,1,1,1,1,1,1,.1))

bisq_mod = data.frame(model = c("Inital weights = 1", "Intial weights = 1 and 0.1"),
                      intercepts = c(bisquare$coef[1], bisquare2$coef[1]),
                      slopes = c(bisquare$coef[2], bisquare2$coef[2]))

gf_point(y ~ x) %>%
  gf_abline(intercept =~ intercepts, slope =~ slopes,
            color =~ model, lty =~ model, lwd = 1.2,
            data = bisq_mod)
```

