---
title: "Homework 3 R markdown"
author: "Abra Brisbin"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_height: 4
    fig_width: 4.5
  pdf_document:
    fig_height: 4
    fig_width: 4.5
  word_document:
    fig_height: 4
    fig_width: 4.5
---


```{r, setup, include=FALSE}
#require(mosaic)   # Load additional packages here 

# Some customization.  You can alter or delete as desired (if you know what you are doing).
#trellis.par.set(theme=theme.mosaic()) # change default color scheme for lattice
knitr::opts_chunk$set(
  echo = TRUE,
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```

Load packages here.
```{r, message=FALSE}

```

#### **Intellectual Property:**  
These problems are the intellectual property of the instructors and may not be reproduced outside of this course.



## Problem 1:  Linear Regression  


In this problem, you will use multiple linear regression to model the incomes of people from Wisconsin.

Data file (on Canvas): *Wisconsin_income.csv*  

Data dictionary (on Canvas): *Wisconsin_income data dictionary.txt*

Public Use Microdata from American Community Survey.  Accessed from http://www2.census.gov/programs-surveys/acs/data/pums/2014/1-Year/ on 27 July 2016.
 




### Question 1 (2 points)
Read in the data Wisconsin_income.csv.  Open the data dictionary in a text editor.  

Notice that the following 8 variables are categorical, but are coded as numbers:    

* Citizenship  
* Class of worker  
* Language spoken at home  
* Marital status  
* Sex  
* Disability  
* Race  
* Hispanic  

Tell R to treat them as factors.  Enter your R code below.


**Code Answer**: 
```{r}

```



### Question 2 (2 points)
Make histograms of people’s total earnings, usual hours worked per week, and travel time to work.  Which of these 3 variables are likely to benefit from log-transformation?  Apply the transformation if appropriate, and enter your R code below.


**Code Answer**: 
```{r}

```



### Question 3 (2 points)
Use *regsubsets()* to perform best subset selection for a linear model for total earnings as a function of all other variables in the data set.  
If you log-transformed any variables in the previous question, use the **transformed** variables,  *not*  the original variables, here.  Consider *all* models with up to 39 variables.  Make a plot summarizing which variables are included in the best model of each size.  Use **Insert** -> **Image** to upload your graph to this question on Canvas.


**Code Answer**: 
```{r}

```



### Question 4 (2 points)
Plot adjusted $R^2$ as a function of number of variables.  Use **Insert** -> **Image** to upload your graph to this question on Canvas.

**Code Answer**: 
```{r}

```



### Question 5 (1 points)
How many variables (not counting the intercept) are in the best model, as measured by adjusted $R^2$?

**Numeric Answer**   (AUTOGRADED on Canvas): 
```{r}

```



### Question 6 (1 points)
How many variables (not counting the intercept) are in the best model, as measured by BIC?

**Numeric Answer**   (AUTOGRADED on Canvas): 
```{r}

```



### Question 7 (4 points)
Set the random seed equal to 3.

Perform 10-fold cross-validation to choose the best size of model (from 1 to 39 variables) based on cross-validation MSE.  Record the mean squared error within each fold for each size of variable.  **Note**: This step will probably take a few minutes to run!  
Enter your R code below.

**Code Answer**: 
```{r}


```

```{r}


```



### Question 8 (1 points)
Find the mean of the MSEs from all the folds with the same number of variables.  Which number of variables gives the lowest cross-validation MSE?

```{r}


```

**Numeric Answer**   (AUTOGRADED on Canvas): 



### Question 9 (2 points)
Estimate the standard error of the cross-validation errors and find the most parsimonious model with a CV error within 1 standard error of the lowest.  How many predictor variables are in the most parsimonious model with a CV error within 1 standard error of the lowest?

**Numeric Answer**   (AUTOGRADED on Canvas): 

```{r}

```



### Question 10 (4 points)
Use `regsubsets` to find the best model for the whole data set which has the number of variables you found in the previous question.  For each variable included in the model, write a sentence giving a possible explanation for the direction of the association.  Refer to variables in plain English. 

**Note**: It may be helpful to refer to the data dictionary and/or a map of Wisconsin, such as https://en.wikipedia.org/wiki/Wisconsin#/media/File:Wisconsin-counties-map.gif.  

**Example**:  "Being in a union is positively associated with earnings.  A possible reason is that unions' collective bargaining tends to be successful in convincing employers to offer higher wages."

```{r}

```

**Text Answer**: 





## Problem 2:  Logistic Regression  


In this problem, you will use logistic regression to predict whether a car has low or high gas mileage.

### Question 11 (2 points)
Write R code to:  

* Load the **Auto** data set into R.  The data set is in the ISLR library.  
* **Tell R to treat the `origin` variable as a factor.**
* Create a binary variable that equals "high" for cars with gas mileage (`mpg`) greater than or equal to the median and "low" for cars with gas mileage below the median.  Tell R to treat it as a factor.  

* **Note:** For consistency with the KNN model in question 18, we want the "low" category of gas mileage to be treated as the "positive" (or 1) category.


Enter your R code below.

**Code Answer**: 
```{r}


```



### Question 12 (2 points)
Use corrplot to make a correlation plot of the numeric variables in **Auto**.  Use **Insert** -> **Image** to upload your graph to this question on Canvas.

**Graph Answer**:
```{r}

```



### Question 13 (2 points)
We plan to use logistic regression to predict the binary gas mileage variable.  Do you have any concerns about collinearity?  If so, select all of the pairs of variables in the following list that cause you concern.  (You may have concerns about some variables that are not on this list.  That's fine; just answer based on the variables listed here.)

**Multiple-Select Answer** (AUTOGRADED on Canvas): One or more of

- weight and horsepower
- mpg and cylinders
- weight and weight
- horsepower and displacement
- acceleration and year
- None of these pairs cause me concern.

(The answer options on Canvas may appear in a different order.)



### Question 14 (1 point)
Perform logistic regression of the binary gas mileage variable on the other variables in **Auto** (excluding `name` and the untransformed `mpg`).  Enter your R code below.

**Code Answer**: 
```{r}

```



### Question 15 (1 point)
Compute the variance inflation factor for each of the predictor variables in the model.  Which variable(s) are above the threshold indicating strong multicollinearity?

**Note:** If none of the variables are above the threshold, double-check your work in question 11.

```{r}

```

**Multiple-Select Answer** (AUTOGRADED on Canvas): One or more of

- cylinders  
- displacement  
- horsepower 
- weight

(The answer options on Canvas may appear in a different order.)



### Question 16 (4 points)
Remove any variables with VIFs greater than or equal to 10.  Set the random seed equal to 3 and perform 10-fold cross-validation.  In each phase of the cross-validation, fit the logistic model (excluding name, continuous mpg, and the variable(s) you found in the previous question) and predict the probability of high gas mileage for each data point in the validation set.  Store all of the probabilities in a single vector.  

**Note**:  Depending on how you set up the formula in the logistic regression, the predict function may give an error, “Factor name has new levels.”  This is complaining about the fact that there are models of car in the validation set that weren’t included in the training data.  But, it’s not really a problem, because we’re not using name as a predictor variable.  You can create a new data frame that excludes name, or you can update the levels of the name factor in the logistic model, as shown [here](http://stackoverflow.com/questions/22315394/factor-has-new-levels-error-for-variable-im-not-using).

Enter your R code below.

**Code Answer**: 

```{r}

```




### Question 17 (2 points)
Create a ROC curve for this model.  What is its AUC?  Enter your answer to 4 decimal places.

**Numeric Answer**   (AUTOGRADED on Canvas):  

```{r}


```



### Question 18 (2 points)
The file `Homework_3_KNN.csv` (available on Homework 3 on Canvas) contains the data needed to construct a ROC curve for a KNN model of binary gas mileage, using 49 nearest neighbors.  Make **one** graph showing the ROC curve of **both** the logistic regression and KNN models.  Your clients should be able to tell which curve represents which model from the graph (without looking at your code).

Use **Insert** -> **Image** to upload your graph on Canvas.

**Graph Answer**:  

```{r}


```


### Question 19 (1 point)
Write 1-2 sentences comparing the logistic regression and KNN models based on their ROC curves.  

**Text Answer**

