---
title: "Lesson 7 Examples--Decision Trees"
author: "Abra Brisbin"
date: "7/1/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
# To use readr, uncomment lines 16, 23, 45, 86, 141, 152
# and comment out lines 24, 48, 87, 140, 153

#library(readr) 
library(tree)
```

# Can we predict whether an object is rock or metal based on its sonar reading?

```{r}
#sonar = read_csv("Sonar_rock_metal.csv", col_names = FALSE)
sonar = read.csv("Sonar_rock_metal.csv", header = FALSE)
head(sonar)
```
```{r}
dim(sonar)
```





```{r}
set.seed(728)
groups = c(rep(1, 140), rep(2, 68)) # 1 represents the training set
random_groups = sample(groups, 208)

in_train = (random_groups == 1)
```

```{r}
# read_csv names the columns "X" + number
#my_tree = tree(factor(X61) ~ ., data = sonar[in_train, ])

# read.csv names the columns "V" + number
my_tree = tree(factor(V61) ~ ., data = sonar[in_train, ])

summary(my_tree)
```




```{r}
pdf("Sonar.pdf") # If this file is open, this line will give an error

par(mar = c(.1,.1,.1,.1))
plot(my_tree)
text(my_tree, pretty = 0)

dev.off()
```






```{r}
preds = predict(my_tree, newdata = sonar[!in_train, ], type = "class")
  # type = "class" returns predicted response classes
  # type = "vector" returns predicted probabilities
```









```{r}
#conf_mat = table(preds, sonar$X61[!in_train])
conf_mat = table(preds, sonar$V61[!in_train])
conf_mat
```
```{r}
sum(diag(conf_mat)) / 68  # Accuracy
1 - sum(diag(conf_mat))/68 # Error rate
```







## Use 10-fold CV to choose the optimal number of leaves
```{r}
set.seed(728)
sonar.cv = cv.tree(my_tree, FUN = prune.misclass)
  # for regression, use prune.tree
sonar.cv
```

```{r}
plot(sonar.cv)
```





## Extracting the optimal number of leaves
```{r}
min(sonar.cv$dev)
which(sonar.cv$dev == min(sonar.cv$dev))
sonar.cv$size[ which(sonar.cv$dev == min(sonar.cv$dev)) ]
```









```{r}
prune_sonar = prune.misclass(my_tree,
                             best = 4)
plot(prune_sonar)
text(prune_sonar, pretty = 0)
```

```{r}
prune_sonar
#prune_sonar$frame # If using readr, copy this into the console (it may not work in the .Rmd)
```







## Changing parameters about when to split the predictor space
```{r}
#detailed_tree = tree(factor(X61) ~ ., data = sonar,control = tree.control(nobs = 208, mindev = 0, minsize = 2))
detailed_tree = tree(factor(V61) ~ ., data = sonar,control = tree.control(nobs = 208, mindev = 0, minsize = 2))

pdf("Detailed_Sonar.pdf")
plot(detailed_tree)
text(detailed_tree)
dev.off()
```







# Decision trees in caret
In this example, we'll build a decision tree in caret to predict crime rates in Boston neighborhoods.

To make decision trees in `caret`, we'll use the `rpart` package.  (Unfortunately, the `tree` package doesn't currently interface with `caret`.)
```{r, message = FALSE}
library(caret)
library(MASS) # for Boston data
library(rpart) # a different package for making trees

library(rattle) # for fancy tree graph
```

We'll use 5-fold cross-validation to select an optimal value of *cp*, the complexity parameter.  This parameter tells R, "When we can’t find a split that improves the fit by at least this much, stop adding splits."  Therefore, lower cp values result in a more complex tree.

```{r}

set.seed(789)
data_used = Boston

ctrl = trainControl(method = "cv", number = 5)
fit_crim = train(crim ~ ., 
             data = data_used,
             method = "rpart",
             na.action = na.exclude,
             tuneGrid = expand.grid(cp = seq(0, .15, by = .01)),
             trControl = ctrl)

fit_crim 

```

We can use the following code (from the `rattle` package) to make fancy graphs of `rpart` models.  In `caret`, the "finalModel" is the best model that `caret` found (in this case, `cp` = 0.13), fit on the whole data set.  This is the model we're graphing.
```{r}
fancyRpartPlot(fit_crim$finalModel)
```

Here, the interpretation would be:

* The most important variable in predicted crime rate is accessibility to radial highways.  Areas with low accessibility (`rad` < 16) have the lowest predicted crime rate, 0.39.
* For areas that are more accessible by highway, the median home value (`medv`) is important.  Areas where the median home value was less than $11,000 (in 1978) have the highest predicted crime rate, 25.

We can view the details of the tree in the same way as we did for trees created using the `tree` package.  Notice that because this is a regression tree, each node is labeled with the mean crime rate of neighborhoods in that node, rather than a pair of probabilities.
```{r}
fit_crim$finalModel
```
